---
title: "HW3.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this assignment will be analyzing data on the Vegetation
and Environment in Dutch Dune Meadows.

To import the data and read the metadata run the following:

```{r}
library(vegan)
data(dune)
data(dune.env)
?dune
```
# The data dune is already structured as a community matrix, dervied from the dune.env data set


```{r}
dune_mds <- metaMDS(dune)

```

#might need to recode moisture as a numeric value

1. Conduct an indirect ordination on the dune plant community. Specifically,
visually examine a NMDS plot using the bray-curtis distance metric. Below is 
some code to help you develop a potential plot that emphasizes the role of the
environmental variable "Moisture". Describe how you interpret the
graphic. What is the goal of creating such a plot? Does this analysis suggest
any interesting findings with respect to the dune vegetation?

    ```{r}
    plot(dune_mds, type='n')
    text(dune_mds, 'sp', cex=.5)
    # generate vector of colors 
    color_vect = rev(terrain.colors(6))[-1]
    points(dune_mds, 'sites', pch=19, 
           col=color_vect[dune.env$Moisture])
    legend('topright', paste("Moisture =", 1:5, sep=''), 
           col=color_vect, pch=19)
    ```
# The above graph shows the moisture levels different species live in. The plot is a very basic summary of the data, showing that the dune vegetation can grow in varying levels of moisture and that the plants that have the same moisture level seem to live in the similar sites.


2. Carry out a direct ordination using CCA in order to test any potential 
hypotheses that you developed after examining the MDS plot. Specifically,
carry out a test of the entire model (i.e., including all constrained axes)
and also carry out tests at the scale of individual explanatory variables
you included in your model if you included more than one variable. Plot your 
results.

```{r}
dune.env$Moisture <- as.numeric(dune.env$Moisture)
cca_veg <- cca(dune ~ ., data= dune.env)
RsquareAdj(cca_veg, 100)
```
```{r}
anova(cca_veg, permutations = 999)
```
```{r}
anova(cca_veg, by='margin', permutations = 999)
```
```{r}
plot(cca_veg, type='n', scaling=1)
orditorp(cca_veg, display='sp', cex=0.5, scaling=1, col='blue')
text(cca_veg, display='cn', col='green')
```


```{r}
anova(cca_veg, by='margin', permutations = 999)
```

3. Do your two analyses agree with one another or complement one another or do
these two analyses seem to be suggesting different take home messages? Which
analysis do you find to be more useful?

#The two analyses do seem to compliment each other.  The cca shows that moisture is linear in its effect, which seems to compliment the indirect ordination.  The indirect ordination shows that moisture is related to site of the plants.  
